06 - PubliliaFitness - Estimating lifetime fecundity and fitness using formulas from Zink 2003

### 2023-08-26 scripts/Publilia-06-fitness_sensu_Zink.Rmd
Moved Zink fitness analysis to it's own script + saved survorship model fits from script 03 to use in 05 + other minor edits to script 05 to run without error

---
title: "Publilia_2019_CHAPTER-2C-fitness"
author: "micah fletcher"
date: "2/24/2022"
output: html_document
---

# Publilia 2019 Chapter - Script 2C: estimating fitness
This is meant to consolidate previous scripts and provide a readable and reproducible record of the analyses for the Publilia Chapter.

## Lifetime fitness using regression predictions
Custom functions for using survivorship regression predictions to estimate fitness
```{r}
ci_for_fitness <- function(data_in, survNom, survDen, 
                        tenure, treatment, vis, model){
  data <- eval(substitute(
    data_in %>%
      add_column(fit = predict(model, newdata = ., type = 'response'))
    ))
  ## grad the inverse link function
  ilink <- family(model)$linkinv
  ## add fit and se.fit on the **link** scale
  data_fit_link <- eval(substitute(
    data %>% add_column(
  setNames(as_tibble(predict(model, data, se.fit = TRUE)[1:2]),
           c('fit_link','se_link'))) %>%
## create the interval and backtransform
  mutate(fit_resp  = ilink(fit_link),
         conf_upr = ilink(fit_link + (2 * se_link)),
         conf_lwr = ilink(fit_link - (2 * se_link)))
    ))
  return(data_fit_link)
}
```

# Estimating fitness using different methods
There are several outdated versions that I've since deleted
 - “v2”: (survivorship ~ ant_treatment * desL) * proportion of time on eggs
 - “v3”: (survivorship ~ ants_ever * desL) * proportion of time on eggs
 - “v4”: (survivorship ~ ants_ever * desP) * proportion of time on eggs
 - “v5”: (survivorship ~ ants_ever * I(desP^2)) * proportion of time on eggs

The current versions are:
 - v1: raw offspring count * proportion of time on eggs
 - v6: (survivorship ~ ant_treatment * I(desL^2)) * proportion of time on eggs

Not time for dissertation but I could experiment with using a square root function for egg laying (more eggs in the first few days of guarding/visiting) and see how robust these results are to our assumptions about egg laying behavior.
```{r}
est_fitness_prep <- visited_1st_4d_data %>%
  filter(!is.na(initiation_date)) %>%
  group_by(Female_ID) %>%
  mutate(started_by_initiating =
           any(clutch_visited_order==1&clutch_initiated_order==1)) %>%
  ungroup() %>%
  select(location, Female_ID, clutch_visited_order, clutch_initiated_order, 
         desert_rel2h = est_rel2h, guard_duration,
         mom_at_init, started_by_initiating) %>%
  left_join(select(clutch_survivorship_data, location:last_desert, 
                   n_eggs, med_t5eggs, n_hatched, n_survived, DormPlant), 
            by=c("location")) %>%
  left_join(select(focalMb4h_sc2h, Female_ID, location, focalMb4h_sc2h),
            by=c("Female_ID", "location")) %>%
  mutate(des=last_des_rel2h, 
         ants=ifelse(ants_removed,"Removed","Control"),
         vis=`visited in 4 days`) %>%
  left_join(select(ant_counts, location, ever_antsP)) %>%
  separate(location, into = c("DormPlant", "Leaf"), sep="_", remove=FALSE) %>%
  left_join(dplyr::select(moms_plant_data, DormPlant, last_visit_ends_P), by="DormPlant") %>%
  mutate(last_desP_rel2h = last_visit_ends_P-est_hatch,
         ant_treatment=ifelse(ants_removed,"Removed","Control"))

# v6: (survivorship ~ ant_treatment * I(desL^2)) * proportion of time on eggs
est_fitness_hatch_fit_v6 <- 
  ci_for_fitness(data_in=est_fitness_prep, model=des_hatch.desL.ant_treatment.quad.fits$mai.qb) %>%
  mutate(est_hatch_fit_v6 = fit) %>%
  select(-(fit:conf_lwr))
est_fitness_Esurv_fit_v6 <- 
  ci_for_fitness(data_in=est_fitness_hatch_fit_v6, model=des_Esurv.desL.ant_treatment.quad.fits$mai.qb) %>%
  mutate(est_Esurv_fit_v6 = fit) %>%
  select(-(fit:conf_lwr))

est_fitness_clutch <- est_fitness_Esurv_fit_v6 %>%
  mutate(propD_b4h = focalMb4h_sc2h/mDb4h_sc2h,
         est_eggs_clutch = med_t5eggs*propD_b4h,
         est_hatched_clutch_v1 = n_hatched*propD_b4h,
         est_survived_clutch_v1 = n_survived*propD_b4h,
         est_hatched_clutch_v6 = est_eggs_clutch*est_hatch_fit_v6,
         est_survived_clutch_v6 = est_eggs_clutch*est_Esurv_fit_v6) %>%
  group_by(Female_ID) %>%
    mutate(est_eggs_life = sum(est_eggs_clutch, na.rm = T),
           est_hatched_life_v1 = sum(est_hatched_clutch_v1, na.rm = T),
           est_survived_life_v1 = sum(est_survived_clutch_v1, na.rm = T),
           est_hatched_life_v6 = sum(est_hatched_clutch_v6, na.rm = T),
           est_survived_life_v6 = sum(est_survived_clutch_v6, na.rm = T),
           n_clutches_iv = n(),
           n_initiated = sum(!is.na(clutch_initiated_order)),
           n_visited = sum(!is.na(initiation_date)&is.na(clutch_initiated_order)),
           n_feed = sum(is.na(initiation_date)),
           init_visit_life_tactic = ifelse(any(mom_at_init), 
                                           ifelse(any(!mom_at_init), "Both", "Initiated only"),
                                           ifelse(!any(mom_at_init), "Visited only", NA)),
           one_both_life_tactic = ifelse(any(mom_at_init), 
                                           ifelse(any(!mom_at_init), "Both", "One"),
                                           ifelse(!any(mom_at_init), "One", NA))) %>%
  ungroup()

est_fitness_future <- est_fitness_clutch %>%
  filter(clutch_visited_order == 1) %>%
  mutate(est_eggs_future = est_eggs_life - est_eggs_clutch,
         est_hatched_future_v1 = est_hatched_life_v1 - est_hatched_clutch_v1,
         est_hatched_future_v6 = est_hatched_life_v6 - est_hatched_clutch_v6,
         est_survived_future_v1 = est_survived_life_v1 - est_survived_clutch_v1,
         est_survived_future_v6 = est_survived_life_v6 - est_survived_clutch_v6)

est_fitness_strat_cats <- est_fitness_future %>%
  dplyr::select(Female_ID, location, desert_rel2h, last_des_rel2h,
         med_t5eggs, est_eggs_clutch, 
         est_hatched_life_v1, est_hatched_life_v6,
         est_hatched_future_v1, est_hatched_future_v6,
         est_survived_life_v1, est_survived_life_v6,
         est_survived_future_v1, est_survived_future_v6,
         females_removed, ants_removed, clutch_initiated_order,
         n_clutches_iv, n_initiated, n_visited, init_visit_life_tactic, `visited in 4 days`)
```

## Check how I'm calculating fecundity
```{r}
# no need to separate females_removed
est_fitness_clutch %>% 
  filter(!is.na(est_eggs_clutch)) %>%
  ggplot(data=., aes(x=des, y=est_eggs_clutch, color=ants, fill=ants)) +
  geom_point() +
  geom_smooth() + ylim(0,60) +
  facet_wrap(vars(vis))

est_fitness_clutch %>%
  arrange(Female_ID) %>%
  select(Female_ID, location, focalMb4h_sc2h, mDb4h_sc2h, propD_b4h, 
         med_t5eggs, est_eggs_clutch, est_eggs_life) #%>% View()
```

## Check how I'm estimating hatching and survival and how many clutches are recovered
It might be that using the regression model with a continuous desertion date explanatory variable just cannot adequately capture the variation and will always deflate mean and variance in lifetime fitness (# N5). If I want a more accurate representation of the variation, maybe I should use discrete desertion date categories (boxplots) and explicitly include variance in the imputation. But will this lead to strange results when the metric determining eggs contributed is continuous? I don't have time to do this for the dissertation but I could explore it for the paper.
```{r}
# Check how I'm estimating hatching and survival and how many clutches are recovered
est_fitness_clutch %>%
  count(females_removed, !is.na(est_eggs_clutch), !is.na(est_hatched_clutch_v1), !is.na(est_survived_clutch_v1))

est_fitness_clutch %>%
  count(females_removed, !is.na(est_eggs_clutch), !is.na(est_hatched_clutch_v6), !is.na(est_survived_clutch_v6))

# It appears that my modeling is overestimating the number of nymphs hatched for clutches with very low est hatched v1
est_fitness_clutch %>% 
  filter(!is.na(est_hatched_clutch_v1),!is.na(est_hatched_clutch_v6)) %>%
  ggplot(data=., aes(x=est_hatched_clutch_v1, y=est_hatched_clutch_v6, color=sum_vD==0)) +
  geom_abline(slope=1) +
  geom_point()

# v6: Consistency is MUCH worse for survivorship; MANY clutches with no surviving nymphs were predicted to have survived
est_fitness_clutch %>%
  filter(!is.na(est_survived_clutch_v1),!is.na(est_survived_clutch_v6)) %>%
  ggplot(data=., aes(x=est_survived_clutch_v1, y=est_survived_clutch_v6, color=vis)) +
  geom_abline(slope=1) +
  geom_point() + facet_wrap(~females_removed)

est_fitness_clutch %>%
  filter(!is.na(est_survived_clutch_v1),!is.na(est_survived_clutch_v6)) %>%
  ggplot(data=., aes(x=est_survived_clutch_v1, y=est_survived_clutch_v6, color=interaction(vis,ants))) +
  geom_abline(slope=1) +
  geom_point()  + xlim(0,60) + ylim(0,60)

### how well does this work from the perspective of imputation?
# hatching: ok correspondence, the spread is tighter, but range is preserved
est_fitness_clutch %>% 
  filter(!is.na(est_hatched_clutch_v1),!is.na(est_hatched_clutch_v6)) %>%
  ggplot(data=., aes(x=des, y=est_hatched_clutch_v1, color=ants, fill=ants)) +
  geom_point() +
  geom_smooth() + ylim(0,60) +
  facet_wrap(vars(vis))

est_fitness_clutch %>% 
  filter(!is.na(est_hatched_clutch_v6)) %>%
  ggplot(data=., aes(x=des, y=est_hatched_clutch_v6, color=ants, fill=ants)) +
  geom_point() +
  geom_smooth() + ylim(0,60) +
  facet_wrap(vars(vis))

# survival: ok correspondence, the spread is MUCH tighter and range is more limited; losing many higher values
est_fitness_clutch %>% 
  filter(!is.na(est_survived_clutch_v1),!is.na(est_survived_clutch_v6)) %>%
  ggplot(data=., aes(x=des, y=est_survived_clutch_v1, color=ants, fill=ants)) +
  geom_point() +
  geom_smooth() + ylim(0,60) +
  facet_wrap(vars(vis))

est_fitness_clutch %>% 
  filter(!is.na(est_survived_clutch_v6)) %>%
  ggplot(data=., aes(x=des, y=est_survived_clutch_v6, color=ants, fill=ants)) +
  geom_point() +
  geom_smooth() + ylim(0,60) +
  facet_wrap(vars(vis))
```

Only females that laid more than one clutch
```{r}
est_fitness_clutch %>%
  filter(!is.na(est_hatched_clutch_v1),!is.na(est_hatched_clutch_v6)) %>%
  ggplot(data=., aes(x=est_hatched_clutch_v1, y=est_hatched_clutch_v6)) +
  geom_abline(slope=1) +
  geom_point() +
  facet_wrap(~n_initiated>1)

est_fitness_clutch %>%
  filter(!is.na(est_survived_clutch_v1),!is.na(est_survived_clutch_v6)) %>%
  ggplot(data=., aes(x=est_survived_clutch_v1, y=est_survived_clutch_v6)) +
  geom_abline(slope=1) +
  geom_point() +
  facet_wrap(~n_initiated>1)
```

### Life strategies: start by deserting early, guarding eggs, guarding nymphs or visited an existing clutch
```{r}
est_fitness_life_strats <- est_fitness_clutch %>%
  group_by(Female_ID) %>%
  # clutch visited order == 0 means she visited before initiation, so it shouldn't count
  mutate(started_by =
           ifelse(any(clutch_visited_order==1&!is.na(clutch_initiated_order)&clutch_initiated_order==1),
                  "initiating",
                  ifelse(any(clutch_visited_order==1&is.na(clutch_initiated_order)|
                             clutch_visited_order==1&!is.na(clutch_initiated_order)&clutch_initiated_order!=1),
                         "visiting", "other")),
         n_clutches_with_eggs_iv = sum(!is.na(clutch_visited_order)&(clutch_visited_order>0), na.rm=T)) %>%
  ungroup() %>%
  filter(clutch_visited_order==1) %>%
  # mutate(desert_cat = ifelse(started_by=="initiating",
  #                                ifelse(desert_rel2h <= -16, "desert",
  #                                       ifelse(desert_rel2h > -16 & desert_rel2h <= 4, "eggs", 
  #                                              ifelse(desert_rel2h >= 4,"nymphs",
  #                                                     NA))),
  #                                ifelse(started_by=="visiting", "visit", 
  #                                       NA)),
  mutate(desert_cat = ifelse(started_by=="initiating",
                                 ifelse(guard_duration <= 10, "desert",
                                        ifelse(guard_duration > 10 & desert_rel2h <= 4, "eggs", 
                                               ifelse(desert_rel2h >= 4,"nymphs",
                                                      NA))),
                                 ifelse(started_by=="visiting", "visit", 
                                        NA)),
         first_clutch_visited_in4d = ifelse(started_by=="initiating",
                                            ifelse(`visited in 4 days` == "Visited", "vis", "not"),
                                            NA),
         est_eggs_future = est_eggs_life - est_eggs_clutch,
         est_hatched_future_v1 = est_hatched_life_v1 - est_hatched_clutch_v1,
         est_hatched_future_v6 = est_hatched_life_v6 - est_hatched_clutch_v6,
         est_survived_future_v1 = est_survived_life_v1 - est_survived_clutch_v1,
         est_survived_future_v6 = est_survived_life_v6 - est_survived_clutch_v6) %>%
  distinct(Female_ID, ants_removed, females_removed, `visited in 4 days`,
           desert_cat, desert_rel2h, guard_duration,
           started_by, init_visit_life_tactic, one_both_life_tactic,
           first_clutch_visited_in4d, n_clutches_with_eggs_iv,
           est_eggs_clutch, est_eggs_life, est_eggs_future, 
           est_hatched_life_v1, est_hatched_life_v6, 
           est_hatched_future_v1, est_hatched_future_v6,
           est_hatched_clutch_v1, est_hatched_clutch_v6,
           est_survived_life_v1, est_survived_life_v6, 
           est_survived_future_v1, est_survived_future_v6,
           est_survived_clutch_v1, est_survived_clutch_v6) %>%
  # remove 4 females that visited clutches with unknown initiation dates
  filter(!is.na(ants_removed)) %>%
  # *** I need to figure out why these are NA!!! ***
  filter(!is.na(est_hatched_clutch_v6))
```

```{r}
est_fitness_life_strats %>%
  filter(!females_removed, !is.na(desert_cat)) %>%
  ggplot(data=., aes(x=desert_cat, y=est_eggs_life, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, outlier.shape=NA) +
  geom_jitter(pch = 21, size=3, width=0.35) +
  ylab("Estimated lifetime fecundity (eggs)")  +
  xlab("Reproductive strategy for first clutch") +
  theme_publilia() +
  facet_wrap(vars(n_clutches_with_eggs_iv>1, ants_removed), nrow=1)

est_fitness_life_strats %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=est_hatched_life_v1, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, outlier.shape=NA) +
  geom_jitter(pch = 21, size=3, width=0.35) +
  ylab("Estimated lifetime fitness (eggs hatched)")  + ylim(0,100) +
  xlab("Reproductive strategy for first clutch") +
  theme_publilia() +
  facet_wrap(vars(n_clutches_with_eggs_iv>1, ants_removed), nrow=1)

est_fitness_life_strats %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=est_hatched_life_v6, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, outlier.shape=NA) +
  geom_jitter(pch = 21, size=3, width=0.35) +
  ylab("Estimated lifetime fitness (eggs hatched)")  + ylim(0,100) +
  xlab("Reproductive strategy for first clutch") +
  theme_publilia() +
  facet_wrap(vars(n_clutches_with_eggs_iv>1, ants_removed), nrow=1)

est_fitness_life_strats %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=est_survived_life_v1, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, outlier.shape=NA) +
  geom_jitter(pch = 21, size=3, width=0.35) +
  ylab("Estimated lifetime fitness (nymphs survived)")  + ylim(0,60) +
  xlab("Reproductive strategy for first clutch") +
  theme_publilia() +
  facet_wrap(vars(n_clutches_with_eggs_iv>1, ants_removed), nrow=1)

est_fitness_life_strats %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=est_survived_life_v6, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, outlier.shape=NA) +
  geom_jitter(pch = 21, size=3, width=0.35) +
  ylab("Estimated lifetime fitness (nymphs survived)")  + ylim(0,60) +
  xlab("Reproductive strategy for first clutch") +
  theme_publilia() +
  facet_wrap(vars(n_clutches_with_eggs_iv>1, ants_removed), nrow=1)
```

## Boxplots

with jitter (messy)
```{r}
est_fitness_life_strats_eggs.p <- est_fitness_life_strats %>%
  filter(!females_removed, !is.na(desert_cat)) %>%
  ggplot(data=., aes(x=desert_cat, y=est_eggs_life, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, outlier.shape=NA) +
  geom_point(pch = 21, position=position_jitterdodge(jitter.width=0.25), size=3) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, outlier.shape=NA, alpha=0) +
  #stat_summary(fun=mean, geom="point", shape=4, size=3) +
  ylab("Lifetime fecundity\n(total eggs laid)")  +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "gray80")) +
  theme_publilia() +
  facet_wrap(~ants_removed, nrow=1) +
  theme(strip.background = element_blank(),strip.text.x = element_blank(),
        axis.title.x = element_blank(),axis.text.x=element_blank(),
        legend.position='none')

est_fitness_life_strats_hatched.p <- est_fitness_life_strats %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=est_hatched_life_v6, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, outlier.shape=NA, key_glyph=draw_key_rect) +
  geom_point(pch = 21, position=position_jitterdodge(jitter.width=0.25), size=3, key_glyph=draw_key_rect) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, outlier.shape=NA, alpha=0, key_glyph=draw_key_rect) +
  #stat_summary(fun=mean, geom="point", shape=4, size=3,  key_glyph=draw_key_rect) +
  ylab("Lifetime fitness\n(total eggs hatched)") +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "gray80")) +
  theme_publilia() +
  facet_wrap(~ants_removed, nrow=1) +
  theme(strip.background = element_blank(),strip.text.x = element_blank(),
        axis.title.x = element_blank(),axis.text.x=element_blank(),
        legend.position = c(0.7,0.9))

est_fitness_life_strats_survived.p <- est_fitness_life_strats %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=est_survived_life_v6, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, outlier.shape=NA) +
  geom_point(pch = 21, position=position_jitterdodge(jitter.width=0.25), size=3) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, outlier.shape=NA, alpha=0) +
  #stat_summary(fun=mean, geom="point", shape=4, size=3) +
  ylab("Lifetime fitness\n(total nymphs survived)") +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "gray80")) +
  theme_publilia() +
  facet_wrap(~ants_removed, nrow=1) +
  theme(strip.background = element_blank(),strip.text.x = element_blank(),
        legend.position='none')

#library(egg)
egg::ggarrange(est_fitness_life_strats_eggs.p,
          est_fitness_life_strats_hatched.p,
          est_fitness_life_strats_survived.p,
          ncol=1, nrow=3)
```

with mean (clean)
```{r}
est_fitness_life_strats_eggs.p <- est_fitness_life_strats %>%
  filter(!females_removed, !is.na(desert_cat)) %>%
  ggplot(data=., aes(x=desert_cat, y=est_eggs_life, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5) +
  stat_summary(fun=mean, geom="point", shape=4, size=3) +
  ylab("Lifetime fecundity\n(total eggs laid)")  +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "gray80")) +
  theme_publilia() +
  facet_wrap(~ants_removed, nrow=1) +
  theme(strip.background = element_blank(),strip.text.x = element_blank(),
        axis.title.x = element_blank(),axis.text.x=element_blank(),
        legend.position = "none")

est_fitness_life_strats_hatched.p <- est_fitness_life_strats %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=est_hatched_life_v6, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, key_glyph=draw_key_rect) +
  stat_summary(fun=mean, geom="point", shape=4, size=3,  key_glyph=draw_key_rect) +
  ylab("Lifetime fitness\n(total eggs hatched)") +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "gray80")) +
  theme_publilia() +
  facet_wrap(~ants_removed, nrow=1) +
  theme(strip.background = element_blank(),strip.text.x = element_blank(),
        axis.title.x = element_blank(),axis.text.x=element_blank(),
        legend.position = c(0.7,0.9))

est_fitness_life_strats_survived.p <- est_fitness_life_strats %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=est_survived_life_v6, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5) +
  stat_summary(fun=mean, geom="point", shape=4, size=3) +
  ylab("Lifetime fitness\n(total nymphs survived)") +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "gray80")) +
  theme_publilia() +
  facet_wrap(~ants_removed, nrow=1) +
  theme(strip.background = element_blank(),strip.text.x = element_blank(),
        legend.position='none')

#library(egg)
egg::ggarrange(est_fitness_life_strats_eggs.p,
          est_fitness_life_strats_hatched.p,
          est_fitness_life_strats_survived.p,
          ncol=1, nrow=3)
```

splitting semelparous and iteroparous females (clean)
```{r}
est_fitness_life_strats_eggs_semel.p <- est_fitness_life_strats %>%
  filter(!females_removed, !is.na(desert_cat), n_clutches_with_eggs_iv==1) %>%
  ggplot(data=., aes(x=desert_cat, y=est_eggs_life, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, key_glyph=draw_key_rect) +
  stat_summary(fun=mean, geom="point", shape=4, size=3,  key_glyph=draw_key_rect) +
  ylab("Lifetime fecundity\n(total eggs laid)")  + ylim(0,160) +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "gray80")) +
  theme_publilia() +
  facet_wrap(~ants_removed, nrow=1) +
  theme(strip.background = element_blank(), strip.text.x = element_blank(),
        axis.title.x = element_blank(),axis.text.x=element_blank(),
        legend.position="none")

est_fitness_life_strats_eggs_itero.p <- est_fitness_life_strats %>%
  filter(!females_removed, !is.na(desert_cat), n_clutches_with_eggs_iv>1) %>%
  ggplot(data=., aes(x=desert_cat, y=est_eggs_life, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, key_glyph=draw_key_rect) +
  stat_summary(fun=mean, geom="point", shape=4, size=3,  key_glyph=draw_key_rect) +
  ylab("Lifetime fecundity\n(total eggs laid)")  + ylim(0,160) +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "gray80")) +
  theme_publilia() +
  facet_wrap(vars(n_clutches_with_eggs_iv>1, ants_removed), nrow=1) +
  theme(strip.background = element_blank(), strip.text.x = element_blank(),
        axis.title.x = element_blank(),axis.text.x=element_blank(),
        legend.position="none")

est_fitness_life_strats_hatched_semel.p <- est_fitness_life_strats %>%
  filter(!females_removed, !is.na(desert_cat), n_clutches_with_eggs_iv==1) %>%
  ggplot(data=., aes(x=desert_cat, y=est_hatched_life_v6, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, key_glyph=draw_key_rect) +
  stat_summary(fun=mean, geom="point", shape=4, size=3,  key_glyph=draw_key_rect) +
  ylab("Lifetime fitness\n(total eggs hatched)") + ylim(0,100) +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "gray80")) +
  theme_publilia() +
  facet_wrap(~ants_removed, nrow=1) +
  theme(strip.background = element_blank(), strip.text.x = element_blank(),
        axis.title.x = element_blank(),axis.text.x=element_blank(),
        legend.position=c(0.75,0.9))

est_fitness_life_strats_hatched_itero.p <- est_fitness_life_strats %>%
  filter(!females_removed, !is.na(desert_cat), n_clutches_with_eggs_iv>1) %>%
  ggplot(data=., aes(x=desert_cat, y=est_hatched_life_v6, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, key_glyph=draw_key_rect) +
  stat_summary(fun=mean, geom="point", shape=4, size=3,  key_glyph=draw_key_rect) +
   ylab("Lifetime fitness\n(total eggs hatched)") + ylim(0,100) +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "gray80")) +
  theme_publilia() +
  facet_wrap(vars(n_clutches_with_eggs_iv>1, ants_removed), nrow=1) +
  theme(strip.background = element_blank(), strip.text.x = element_blank(),
        axis.title.x = element_blank(),axis.text.x=element_blank(),
        legend.position=c(0.75,0.9))

est_fitness_life_strats_survived_semel.p <- est_fitness_life_strats %>%
  filter(!females_removed, !is.na(desert_cat), n_clutches_with_eggs_iv==1) %>%
  ggplot(data=., aes(x=desert_cat, y=est_survived_life_v6, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, key_glyph=draw_key_rect) +
  stat_summary(fun=mean, geom="point", shape=4, size=3,  key_glyph=draw_key_rect) +
  ylab("Lifetime fitness\n(total nymphs survived)") + ylim(0,30) +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "gray80")) +
  theme_publilia() +
  facet_wrap(~ants_removed, nrow=1) +
  theme(strip.background = element_blank(), strip.text.x = element_blank(),
        legend.position="none",
        axis.text.x=element_text(angle=45, vjust=1, hjust=1))

est_fitness_life_strats_survived_itero.p <- est_fitness_life_strats %>%
  filter(!females_removed, !is.na(desert_cat), n_clutches_with_eggs_iv>1) %>%
  ggplot(data=., aes(x=desert_cat, y=est_survived_life_v6, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, key_glyph=draw_key_rect) +
  stat_summary(fun=mean, geom="point", shape=4, size=3,  key_glyph=draw_key_rect) +
  ylab("Lifetime fitness\n(total nymphs survived)") + ylim(0,30) +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "gray80")) +
  theme_publilia() +
  facet_wrap(vars(n_clutches_with_eggs_iv>1, ants_removed), nrow=1) +
  theme(strip.background = element_blank(), strip.text.x = element_blank(),
        legend.position="none",
        axis.text.x=element_text(angle=45, vjust=1, hjust=1))

egg::ggarrange(est_fitness_life_strats_eggs_semel.p,est_fitness_life_strats_eggs_itero.p,
          est_fitness_life_strats_hatched_semel.p,est_fitness_life_strats_hatched_itero.p,
          est_fitness_life_strats_survived_semel.p,est_fitness_life_strats_survived_itero.p,
          ncol=2, nrow=3)
```

Should I include this raw data in the supplement? I think so.
```{r}
est_fitness_life_strats %>%
  filter(!females_removed, !is.na(desert_cat)) %>%
  ggplot(data=., aes(x=desert_cat, y=est_eggs_life, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, outlier.shape=NA) +
  geom_point(pch = 21, position=position_jitterdodge(jitter.width=0.25), size=3) +
  ylab("Estimated lifetime fecundity (eggs)")  +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "white")) +
  theme_publilia() +
  facet_wrap(vars(ants_removed), nrow=1)

est_fitness_life_strats %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=est_hatched_life_v1, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, outlier.shape=NA) +
  geom_point(pch = 21, position=position_jitterdodge(jitter.width=0.25), size=3) +
  ylab("Estimated lifetime fitness (eggs hatched)") +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "white")) +
  theme_publilia() +
  facet_wrap(vars(ants_removed), nrow=1)

est_fitness_life_strats %>%
  filter(!females_removed) %>%
  ggplot(data=., aes(x=desert_cat, y=est_survived_life_v1, fill=ants_removed)) +
  geom_boxplot(position = position_dodge(0.75), size = 1.5, outlier.shape=NA) +
  geom_point(pch = 21, position=position_jitterdodge(jitter.width=0.25), size=3) +
  ylab("Estimated lifetime fitness (nymphs survived)") +
  xlab("Reproductive strategy for first clutch") +
  scale_fill_manual(labels=c("Control","Ants removed"), values=c("brown1", "white")) +
  theme_publilia() +
  facet_wrap(vars(ants_removed), nrow=1)
```

## Split violin plots
geom_split_violin
```{r}
GeomSplitViolin <- ggproto("GeomSplitViolin", GeomViolin, 
                           draw_group = function(self, data, ..., draw_quantiles = NULL) {
  data <- transform(data, xminv = x - violinwidth * (x - xmin), xmaxv = x + violinwidth * (xmax - x))
  grp <- data[1, "group"]
  newdata <- plyr::arrange(transform(data, x = if (grp %% 2 == 1) xminv else xmaxv), if (grp %% 2 == 1) y else -y)
  newdata <- rbind(newdata[1, ], newdata, newdata[nrow(newdata), ], newdata[1, ])
  newdata[c(1, nrow(newdata) - 1, nrow(newdata)), "x"] <- round(newdata[1, "x"])

  if (length(draw_quantiles) > 0 & !scales::zero_range(range(data$y))) {
    stopifnot(all(draw_quantiles >= 0), all(draw_quantiles <=
      1))
    quantiles <- ggplot2:::create_quantile_segment_frame(data, draw_quantiles)
    aesthetics <- data[rep(1, nrow(quantiles)), setdiff(names(data), c("x", "y")), drop = FALSE]
    aesthetics$alpha <- rep(1, nrow(quantiles))
    both <- cbind(quantiles, aesthetics)
    quantile_grob <- GeomPath$draw_panel(both, ...)
    ggplot2:::ggname("geom_split_violin", grid::grobTree(GeomPolygon$draw_panel(newdata, ...), quantile_grob))
  }
  else {
    ggplot2:::ggname("geom_split_violin", GeomPolygon$draw_panel(newdata, ...))
  }
})

geom_split_violin <- function(mapping = NULL, data = NULL, stat = "ydensity", position = "identity", ..., 
                              draw_quantiles = NULL, trim = TRUE, scale = "area", na.rm = FALSE, 
                              show.legend = NA, inherit.aes = TRUE) {
  layer(data = data, mapping = mapping, stat = stat, geom = GeomSplitViolin, 
        position = position, show.legend = show.legend, inherit.aes = inherit.aes, 
        params = list(trim = trim, scale = scale, draw_quantiles = draw_quantiles, na.rm = na.rm, ...))
}
```

>>>>

---
title: '06 - PubliliaFitness - Estimating lifetime fecundity and fitness using formulas from Zink 2003'
author: "Micah Fletcher"
date: '`r Sys.Date()`'
output: 
  rmdformats::robobook:
  fig_width: 9
toc_depth: 3
---

# Load libraries

```{r}
library(multcompView)
source(here("R/Publilia-05-fitness-functions.R"))
```

# 0 - Data import

```{r}
surv_reg_data <- read_rds(here("saved_rds/surv_reg_data.rds"))
```
